---
title: 
author: 
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(PASTRI)
library(tidyverse)
library(parallel)
library(ape)
library(ggtree)
library(argparse)
library(magrittr)
library(patchwork)
library(ggridges)
library(reshape2)
library(tidyr)
library(stats)
library(Matrix)
library(expm)
library(pracma)
library(ggplot2)
library(data.table)
```

1. Get the normalized depth of tip (cell) pairs

In this section, we use the function calculate_lca_depths() to compute the normalized depth for each node (including tip/cell pairs) from the specified tree file (sample.nwk)
```{r}
# Assume that calculate_lca_depths() is a custom or externally provided function
# that takes an NWK file path and returns a data structure (e.g., data frame)
# containing the depth of each node/tip.

tree_file <- system.file("extdata", "HSC.nwk", package = "PASTRI")

sample.node.depth <- calculate_lca_depths(tree.path = tree_file)

# Display the first few rows of the resulting node depth information
head(sample.node.depth)
```

2. Get the optimal transition matrix

Next, we demonstrate the use of the function get_optimal_transition_matrix() to obtain the optimal transition matrix. We will illustrate multiple iterations over a set of depth thresholds (depth) and combine all results into a single data frame df_mrca.mheight.

Analysis was restricted to single-cell leaves from the lineage tree (i.e., cellNum == 1 in "data/HSC_nodeInfos.csv").
```{r}

# Read the CSV file that contains cell information
cell_file <- system.file("extdata", "HSC_nodeInfos.csv", package = "PASTRI")
cell <- read.csv(cell_file, stringsAsFactors = FALSE)


# Specify a series of depth thresholds to test
depth <- c(0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0)

# Set cell type
cell_type <- c("C4", "C5", "C6", "C7", "C8")

# Constrained matrix
# represent prior knowledge such as irreversible state transitions
Bound_matrix <- matrix(Inf, nrow = length(cell_type), ncol = length(cell_type))
Bound_matrix[upper.tri(Bound_matrix)] <- 0


# Use lapply to loop over each depth threshold
df_depth <- lapply(seq(length(depth)), function(d) {
  sel.depth <- depth[d]
  print(sel.depth)  # Print the current depth threshold
    
  # Call get_optimal_transition_matrix() with the specified parameters
  mrca.mheight_optimal_results_d <- get_optimal_transition_matrix(
    node_pair_depth = sample.node.depth,
    cell_info = cell,
    Sel_u = "lca_normalized_height",
    fi_depth = sel.depth,
    Bound_Matrix = Bound_matrix,
    cell_type_list = cell_type
    )
  return( mrca.mheight_optimal_results_d)
})

# Show the transition matrix of depth 0.2
df_depth[[4]]$optimal_matrix %>% t()
```
